Towards Verifying the Bitcoin-S Library

### 2019-06-26 paper/paper.tex
remove needless transformation

  The code we use for the following sections is in reduced folder.
}

@@

\paragraph{Transforming the Code}

@@

@@

\subsection{Writing Specification for the Property}

Now we will look into the previous error.

\subsection{Propagating Require}

There is another problem with Bitcoin-S.
Bitcoin-S-Core uses require as a fail-fast method whereas Stainless needs it to verify the code.

Stainless output now:
\begin{lstlisting}[style=stainless]
[  Info  ]  - Checking cache: 'postcondition' VC for + @12:3...
[  Info  ]  - Checking cache: 'adt invariant' VC for apply @34:39...
[  Info  ]  - Checking cache: 'adt invariant' VC for apply @62:38...
[  Info  ] Cache miss: 'adt invariant' VC for apply @62:38...
[  Info  ] Cache hit: 'adt invariant' VC for apply @34:39...
[  Info  ] Cache hit: 'postcondition' VC for + @12:3...
[  Info  ]  - Now solving 'adt invariant' VC for apply @62:38...
[  Info  ]  - Result for 'adt invariant' VC for apply @62:38:
[Warning ]  => INVALID
[Warning ] Found counter-example:
[Warning ]   bigInt: BigInt -> -9223372036854775809
[  Info  ]  - Checking cache: 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36...
[  Info  ] Cache miss: 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36...
[  Info  ]  - Now solving 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36...
[  Info  ]  - Result for 'precond. (call checkResult(thiss, underlying(thiss) + u ...)' VC for + @17:36:
[Warning ]  => INVALID
[Warning ] Found counter-example:
[Warning ]   num: Number   -> Int64Impl(9223372036854775807)
[Warning ]   thiss: Number -> Int64Impl(1)
\end{lstlisting}

Corresponding code:
\begin{lstlisting}[style=scala]
sealed abstract class Number {
  /** The underlying scala number used to to hold the number */
  protected def underlying: BigInt

  /**
    * Checks if the given result is within the range
    * of this number type
    */
  private def checkResult(result: BigInt): BigInt = {
    require(
         result <= BigInt("9223372036854775807")
      && result >= BigInt("-9223372036854775808"))
    result
  }
}
\end{lstlisting}

But how does Stainless find a counter example ignoring the require in checkResult?
Since Stainless is a static verification tool, it tests every possibility.
So it can use a number bigger than the maximum Int64 and pass it to the addition.
The require in checkResult fails.

Thus, we need to add the restriction from checkResult to the addition too.
\begin{lstlisting}[style=scala]
override def +(num: Int64): Int64 = {
  require(
       num.underlying <= BigInt("9223372036854775807")
    && num.underlying >= BigInt("-9223372036854775808")
    && this.underlying <= BigInt("9223372036854775807")
    && this.underlying >= BigInt("-9223372036854775808")
  )
  apply(checkResult(underlying + num.underlying))
}
\end{lstlisting}

Stainless finds another counter example:
\begin{lstlisting}[style=stainless]
[Warning ] Found counter-example:
[Warning ]   num: { x: Object | @unchecked isInt64(x) }    -> Int64Impl(1)
[Warning ]   thiss: { x: Object | @unchecked isNumber(x) } ->
                Int64Impl(9223372036854775807)
\end{lstlisting}

Sure, when adding one to the maximum Int64 the require does not hold anymore.
Since we do only allow zero as a parameter, the easiest way is to restrict it to zero here too.

Code after:
\begin{lstlisting}[style=scala]
sealed abstract class Number {
  /** The underlying scala number used to to hold the number */
  protected def underlying: BigInt

  def +(num: Int64): Int64 = {
    require(
          num == Int64.zero
      && this.underlying <= BigInt("9223372036854775807")
      && this.underlying >= BigInt("-9223372036854775808"))
    apply(checkResult(underlying + num.underlying))
  }

  /**
    * Checks if the given result is within the range
    * of this number type
    */
  private def checkResult(result: BigInt): BigInt = {
    require(
          result <= BigInt("9223372036854775807")
      && result >= BigInt("-9223372036854775808"))
    result
  }
}
\end{lstlisting}

The same problem occurs in the Int64 constructor.

Before:
\begin{lstlisting}[style=scala]
sealed abstract class Int64 extends SignedNumber {
  override def apply: BigInt => Int64 = Int64(_)
}

case object Int64 extends BaseNumbers[Int64] {
  lazy val zero = Int64(0)
  lazy val one = Int64(1)

private case class Int64Impl(underlying: BigInt) extends Int64 {
  require(underlying >= BigInt("-9223372036854775808"))
  require(underlying <= BigInt("9223372036854775807"))
}
\end{lstlisting}

After:
\begin{lstlisting}[style=scala]
sealed abstract class Int64 extends SignedNumber {
  override def apply: BigInt => Int64 = bigInt => {
    require(
          bigInt >= BigInt("-9223372036854775808")
      && bigInt <= BigInt("9223372036854775807"))

    Int64(bigInt)
  }
}

case object Int64 extends BaseNumbers[Int64] {
  lazy val zero = Int64(0)
  lazy val one = Int64(1)

  def apply(bigInt: BigInt): Int64 = {
    require(
         bigInt >= BigInt("-9223372036854775808")
      && bigInt <= BigInt("9223372036854775807"))

    Int64Impl(bigInt)
  }
}

private case class Int64Impl(underlying: BigInt) extends Int64 {
  require(underlying >= BigInt("-9223372036854775808"))
  require(underlying <= BigInt("9223372036854775807"))
}
\end{lstlisting}

>>>>

  The code we use for the following sections is in reduced folder.

@@

\paragraph{Transforming the Code.}

@@

\paragraph{The Specification.}

@@
